## Metadata
---
title: "NFLFG"
author: "Jackson Perry"
date: "11/8/2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
```{r, include=FALSE, echo=FALSE}
library(tidyverse)
library(caret)
library(glmnet)
library(lubridate)
library(readr)
library(splitstackshape)
library(timechange)
```

## Read csv files
```{r, include=FALSE, echo=FALSE}
plays <-read_csv("plays.csv")
players <-read_csv("players.csv")
games <- read_csv("games.csv")
  games$gameDate <- as.Date(games$gameDate,format="%m/%d/%Y")
  games$gameDateTimeEastern <- as.POSIXct(paste(games$gameDate, games$gameTimeEastern), 
                                   format="%Y-%m-%d %H:%M:%S", "America/New_York")
stadiums <-read_csv("Stadium List.csv")
weather <- read_csv("games_weather.csv")
```

## Understand list of teams, add team numbers to games
```{r, include=FALSE, echo=FALSE}
games$homeTeamAbbr[games$homeTeamAbbr == "OAK"] <- "LV"  # rename OAK to LV (Okaland Raiders moved to Las Vegas)
games$visitorTeamAbbr[games$visitorTeamAbbr == "OAK"] <- "LV"  # rename OAK to LV

# Craete list of team abbreviations with unique ID
teamlist <- data.frame(unique(games$homeTeamAbbr, incomparables = FALSE)) # a list of unique team names
teamlist <- teamlist[order(teamlist),] # sort
teamlist <- data.frame(teamlist) # create dataframe
names(teamlist) <- c("team_abbreviation") #assign name to column
getanID(teamlist, "team_abbreviation") # get an id for each row
teamlist <- teamlist %>% mutate(team_id = row_number()) # get an id for each row
colnames(teamlist) <- c("team_abbreviation","hometeam_ID#") # name each column

games <- merge(games, teamlist, by.x = "homeTeamAbbr", by.y = "team_abbreviation", all=TRUE) # match home team
colnames(teamlist) <- c("team_abbreviation","visitorteam_ID#") # rename for visitor team
games <- merge(games, teamlist, by.x = "visitorTeamAbbr", by.y = "team_abbreviation", all=TRUE) # match visitor team
```



## Convert game time from EST to local time
```{r, include=FALSE, echo=FALSE}
games <- left_join(games, stadiums, by = c("homeTeamAbbr" = "Abbreviation"))
games <- subset(games, select = -c(Team, Conference))

## Account for games played in other venues
webley_games = c(2018101407, 2018102100, 2018102800, 2019102706, 2019110300)
for (i in webley_games) {
  games$Latitude[games$gameId == i] = 51.5560247 # Webley Stadium
  games$Longitude[games$gameId == i] = -0.2818064
  games$ZIP[games$gameId == i] = NA
  games$City[games$gameId == i] = "London"
  games$State[games$gameId == i] = "ENG"
  games$Dome[games$gameId == i] = "Open Air"
  games$`Turf Type`[games$gameId == i] = "Natural"
  games$`Time Zone`[games$gameId == i] = "Europe/London"
}

tottenham_games = c(2019100605, 2019101300)
for (i in tottenham_games) {
  games$Latitude[games$gameId == i] = 51.6042781 # Tottenham Stadium
  games$Longitude[games$gameId == i] = -0.0662422
  games$ZIP[games$gameId == i] = NA
  games$City[games$gameId == i] = "London"
  games$State[games$gameId == i] = "ENG"
  games$Dome[games$gameId == i] = "Open Air"
  games$`Turf Type`[games$gameId == i] = "Natural"
  games$`Time Zone`[games$gameId == i] = "Europe/London"
}

estadio_azteca_games = c(2019111800)
for (i in estadio_azteca_games) {
  games$Latitude[games$gameId == i] = 19.3028607 # Estadio Azteca Stadium
  games$Longitude[games$gameId == i] = -99.1505277
  games$ZIP[games$gameId == i] = NA
  games$City[games$gameId == i] = "Mexico City"
  games$State[games$gameId == i] = "MX"
  games$Dome[games$gameId == i] = "Open Air"
  games$`Turf Type`[games$gameId == i] = "Natural"
  games$`Time Zone`[games$gameId == i] = "America/Mexico_City"
}


games$gameTimeLocal <- NA
  
## Set local times for each area
games$gameTimeLocal[games$`Time Zone` == 'America/Los_Angeles'] <- 
    format(games$gameDateTime, tz="America/Los_Angeles")[games$`Time Zone` == 'America/Los_Angeles']
games$gameTimeLocal[games$`Time Zone` == 'America/New_York'] <- 
    format(games$gameDateTime, tz="America/New_York")[games$`Time Zone` == 'America/New_York']
games$gameTimeLocal[games$`Time Zone` == 'America/Chicago'] <- 
    format(games$gameDateTime, tz="America/Chicago")[games$`Time Zone` == 'America/Chicago']
games$gameTimeLocal[games$`Time Zone` == 'America/Denver'] <- 
    format(games$gameDateTime, tz="America/Denver")[games$`Time Zone` == 'America/Denver']
games$gameTimeLocal[games$`Time Zone` == 'America/Phoenix'] <- 
    format(games$gameDateTime, tz="America/Phoenix")[games$`Time Zone` == 'America/Phoenix']
games$gameTimeLocal[games$`Time Zone` == 'America/Indiana/Indianapolis'] <- 
    format(games$gameDateTime, tz="America/Indiana/Indianapolis")[games$`Time Zone` == 'America/Indiana/Indianapolis']
games$gameTimeLocal[games$`Time Zone` == 'Europe/London'] <- 
    format(games$gameDateTime, tz="Europe/London")[games$`Time Zone` == 'Europe/London']
games$gameTimeLocal[games$`Time Zone` == 'America/Mexico_City'] <- 
    format(games$gameDateTime, tz="America/Mexico_City")[games$`Time Zone` == 'America/Mexico_City']

```

## Join Tables and Rename
```{r, include=FALSE, echo=FALSE}
nflData1 <- inner_join(plays, players, by = c("kickerId" = "nflId"))
nflData2 <- inner_join(games, nflData1, by = c("gameId" = "gameId"))
nflData3 <- inner_join(weather, nflData2, by = c("game_id" = "gameId"))
nflData <- nflData3
```

## Filter dataset for relevant columns
```{r, include=FALSE, echo=FALSE}
df <- nflData %>%
  dplyr::select(homeTeamAbbr, visitorTeamAbbr, quarter, possessionTeam, specialTeamsPlayType, specialTeamsResult, preSnapHomeScore, preSnapVisitorScore, kickLength, birthDate, gameDate, gameClock, gameTimeEastern, height, weight, kickerId, Temperature, DewPoint, Humidity, Precipitation, WindSpeed, WindDirection, Pressure, EstimatedCondition)
#Filter for field goals
fgdf <- df %>%
  filter(specialTeamsPlayType == "Field Goal")
```

#Create Home Field Advantage variable
```{r, include=FALSE, echo=FALSE}
#Transform variables
fgdf <- fgdf %>%
  mutate(homeFieldAdv = case_when(homeTeamAbbr == possessionTeam ~ 1, homeTeamAbbr != possessionTeam ~ 0))

#Create Winning & Tied at Time of Attempt variables
#Create temporary variable for dummy variable calculations
fgdf <- fgdf %>%
  mutate(homeUp = case_when(preSnapHomeScore > preSnapVisitorScore ~ 1,   preSnapHomeScore < preSnapVisitorScore ~ 0))

#Create tied variable
 fgdf <- fgdf %>%
   mutate(tied = case_when(preSnapHomeScore == preSnapVisitorScore ~ 1,   preSnapHomeScore != preSnapVisitorScore ~ 0))
 
#Create winning variable
fgdf <- fgdf %>%
  mutate(winning = case_when(homeFieldAdv <= homeUp ~ 1, 
                             homeFieldAdv >= homeUp ~ 0)) 

#If game is tied at time of attempt, winning is 0
fgdf <- fgdf %>%
  mutate(winning = case_when(tied == 0 ~ 1, 
                             tied == 1 ~ 0))
```

#Create Kicker Age variable
```{r, include=FALSE, echo=FALSE}
#Transform variables
#Convert char to date
fgdf1 <- fgdf
fgdf1 <- fgdf1 %>%
        mutate(birthDate = as.Date(birthDate, tryFormats = c("%m/%d/%Y", "%Y-%m-%d", "$mm/$dd/$yyyy", "$m/$d/$y")))
fgdf1 <- fgdf1 %>%
        mutate(gameDate = as.Date(gameDate, tryFormats = c("%m/%d/%Y", "%Y-%m-%d", "$mm/$dd/$yyyy", "$m/$d/$y")))
#Calculate years between dates for age
fgdf1 <- fgdf1 %>%
        mutate(kickerAge = as.numeric(difftime(birthDate,gameDate),
                                      units = "weeks")/-52.25)
```

#Create Field Goal Completion dummy variable
```{r, include=FALSE, echo=FALSE}
#Transform variables
fgdf <- fgdf %>%
  mutate(fgComplete = case_when(specialTeamsResult == "Kick Attempt Good" ~ 1, 
                             specialTeamsResult != "Kick Attempt Good" ~ 0)) 

```

#Create Quarter dummy variables
```{r, include=FALSE, echo=FALSE}
#Transform variables
fgdf <- fgdf %>%
  mutate(q1 = case_when(quarter == 1 ~ 1, 
                             quarter != 1 ~ 0))
fgdf <- fgdf %>%
  mutate(q2 = case_when(quarter == 2 ~ 1, 
                             quarter != 2 ~ 0))
fgdf <- fgdf %>%
  mutate(q3 = case_when(quarter == 3 ~ 1, 
                             quarter != 3 ~ 0))
fgdf <- fgdf %>%
  mutate(q4 = case_when(quarter == 4 ~ 1, 
                             quarter != 4 ~ 0))
fgdf <- fgdf %>%
  mutate(overtime = case_when(quarter == 5 ~ 1, 
                             quarter != 5 ~ 0))

```

```{r, include=FALSE, echo=FALSE}
#Transform variables
#Create Estimated Condition dummy variables
fgdf <- fgdf %>%
  mutate(clear = case_when(EstimatedCondition == "Clear" ~ 1, 
                             EstimatedCondition != "Clear" ~ 0))
fgdf <- fgdf %>%
  mutate(light_rain = case_when(EstimatedCondition == "Light Rain" ~ 1, 
                             EstimatedCondition != "Light Rain" ~ 0))
fgdf <- fgdf %>%
  mutate(moderate_rain = case_when(EstimatedCondition == "Moderate Rain" ~ 1, 
                             EstimatedCondition != "Moderate Rain" ~ 0))
fgdf <- fgdf %>%
  mutate(heavy_rain = case_when(EstimatedCondition == "Heavy Rain" ~ 1, 
                             EstimatedCondition != "Heavy Rain" ~ 0))
fgdf <- fgdf %>%
  mutate(light_snow = case_when(EstimatedCondition == "Light Snow" ~ 1, 
                             EstimatedCondition != "Light Snow" ~ 0))
fgdf <- fgdf %>%
  mutate(moderate_snow = case_when(EstimatedCondition == "Moderate Snow" ~ 1, 
                             EstimatedCondition != "Moderate Snow" ~ 0))
fgdf <- fgdf %>%
  mutate(heavy_snow = case_when(EstimatedCondition == "Heavy Snow" ~ 1, 
                             EstimatedCondition != "Heavy Snow" ~ 0))

```

```{r, include=FALSE, echo=FALSE}
#Tidy the data
#Delete non-numeric variables
fgdf <- fgdf %>%
  dplyr::select(-visitorTeamAbbr, -homeTeamAbbr, -EstimatedCondition, -quarter, -possessionTeam, -specialTeamsPlayType, -specialTeamsResult, -birthDate, -gameDate, -gameTimeEastern, -height, -kickerId, -homeUp, -gameClock )

#Remove NAs
fgdf <- na.omit(fgdf)
```

```{r, include=FALSE, echo=FALSE}
#Select Predictors
cor(fgdf[,1:28])
```

